Implementing Capon Adaptive Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging

### 2012-10-08 HRUIonGPU/conf_article.tex

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{
Jon Petter \AA{}sen\IEEEauthorrefmark{1},
Jo Inge Buskenes\IEEEauthorrefmark{2},
Carl-Inge C. Nilsen\IEEEauthorrefmark{2},
Andreas Austeng\IEEEauthorrefmark{2},
Sverre Holm\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}
}
\IEEEauthorblockA{
\IEEEauthorrefmark{1} Mi Lab, Norwegian University of Science and Technology, Trondheim, Norway,\\
\IEEEauthorrefmark{2} Department of Informatics, University of Oslo, Oslo, Norway
}
}

@@

For a typical cardiac ultrasound image of $80\times400$ pixels ($70\degree$ sector, 15cm range) acquired using a 2.5MHz, M=64 element phased array, we obtain 10 fps (subarray length L=M/2, temporal smoothing over 3 samples). If we do a 2-element pre-beamforming to get the channel count down to 32, the frame rate increases to 44 fps. For a 32 element phased array we need less beams to cover the sector ($40\times400$ pixels), hence with the same parameters the frame rate increases to 87 fps. After optimization the bottleneck of the Capon algorithm is to find the inverse covariance matrix for all pixels. This is an $O(L^3)$ operation, and when L is larger than 16 (M = 64), real time performance is not possible for cardiac imaging with current high-end hardware. The target GPU has been Nvidia’s Quadro 6000, capable of 1 Tflops.

@@

\begin{figure}
\centerline{\includegraphics[width=3in]{gfx/beamforming_mv_lowres.png}}
\caption{Adaptive beamforming. Impinging signal is demodulated, and aligned in phase by applying steering and focusing delays $\Delta_m[n]$. Adaptive weights, $\vec{w}[n]$, are then calculated based on in-phase data, and finally the output is formed by a weighted coherent sum.}
\label{fig:mvbf}
\end{figure}

@@

Second, we calculate the Capon weight vector as
\begin{align}\label{eq:w}
\vec{w}[n] = \frac{\mat{\hat{R}}[n]^{-1}\vec{a}}{\vec{a}^H\mat{\hat{R}}[n]^{-1}\vec{a}} = \frac{\vec{b}}{\vec{a}^H\vec{b}},
\end{align}
where $\vec{a} = [1_0, 1_1, ..., 1_{L-1}]^T = \vec{1}_L$ when $\vec{x}$ is pre-delayed.

And last, the beamformer output is calculated by combining all $N_L$ subarrays weighted with the same set of adaptive weights,
\begin{align}
z[n] &= \frac{1}{N_L}\vec{w}[n]^H \sum_{l=0}^{N_L-1} \vec{x}_l^{\Delta}[n]. \label{eq:z_mv}
\end{align}

@@

\subsection{GPU Compute Model}
CUDA by Nvidia has been our choice of GPU developer environment \cite{Nvidia2011}. The CUDA compute model is based on execution of a kernel function across a grid of compute threads, where the kernel function describes the work to be done by each thread. As depicted in Figure \ref{fig:gpulayout}, each position in the grid holds a block of threads, with maximal size of 1024 threads, that can be synchronized with little cost if needed. 
%Each block is further divided into groups of 32 threads, known as a warp. The maximum number of threads inside a block is restricted to 512, hence 16 warps. 
From a CUDA perspective the GPU consists of one or more stream multi processors (SMs), where each SM is capable of 32 or more, depending on the architecture, concurrent multiplications and/or additions. A group of blocks is scheduled to each SM until all blocks have been processed.

A SM has available a limited amount of near-core memory and registers. Each thread therefore has to be as light on memory as possible and has to perform enough instructions per used byte in order to hide memory latency. Global reads and writes on the GPU, in addition to CPU-GPU transfers, have high penalty, and should be minimized as much as possible. All of these constraints combined pose a real challenge as the target problem has to map on to the parallel framework both on a grid, block and memory consumption level in order to benefits from parallel acceleration. In the next section, we will describe how this mapping can be done for Capon beamforming of cardiac ultrasound data.

@@

\section{Parallel Capon}\label{sec:meth}
If we analyze the beamformer in (\ref{eq:z_mv}) in combination with (\ref{eq:w}) and (\ref{eq:R}), we see that each calculation of a set of Capon weights is independent across the outputs $z[n]$. As shown in Figure \ref{fig:gpulayout}, a first level of granularity is therefore to divide the image into a grid, where the output in each cell can be calculated independently of all others. Further, based on memory dependencies and to hold memory consumption per thread at a minimum, one should recognise that a group of threads should process one output in the final image. Continuing with the calculation of one weight vector, the algorithm can be broken down into three main steps that, to some extent, can be further divided into a small set of parallel processes. That is, for each data vector $\vec{x}^\Delta$ in the image:
\begin{align*}
%\begin{array}{l}
\text{1) } &\text{Calculate the sample covariance matrix } \mat{\hat{R}} \text{ as in } \text{(\ref{eq:diag})}\\
\text{2) } &\text{Solve the linear system of equations } \vec{b} =\mat{\hat{R}}^{-1}\vec{1}_L\\
\text{3) } &\text{Calculate } \vec{w} = \vec{b}/\vec{1}_{L}^H\vec{b} \text{ and the amplitude as in (\ref{eq:z_mv})}
%\end{array}
\end{align*}

@@

\subsection{Calculating Multiple Sample Covariance Matrices}
The straightforward sum in (\ref{eq:R}) has complexity $O(L^2N_LN_K)$. Note that for a typical array configuration ($M \in [32, 128]$) we do more spatial smoothing than temporal smoothing. Hence, $N_K$ is usually small compared to the number of subarrays ($N_L$). If we let $L = M/\alpha$, $L^2N_L$ becomes $(\alpha - 1)L^3 + L^2$, $\alpha \in [1, M]$. A common choice for $\alpha$ is between two and four (cite!!!). In this interval, calculating the sample covariance matrix is cubic with $L$. This is actually the same as for inverting an $L \times L$ matrix. However little attention has been devoted to this step in the literature.

From (\ref{eq:R}) it is obvious that calculation of one element in $\mat{\hat{R}}$ is independent of all the other elements. A natural level of granularity would then be to assign one thread to each element of $\mat{\hat{R}}$ \cite{Chen2011}. This will minimize global reads and writes needed per thread, but unfortunately the maximum number of threads per block puts an upper limit on the subarray length, $L_{max}=32$, since we want to avoid block-to-block synchronization.

The cubic complexity with $L$ and linear complexity with $N_K$ can be reduced by realizing that the calculation in (\ref{eq:R}) overlaps both across subarrays and in time. Element $(i,j)$ of $\mat{\hat{R}}[n]$ can be calculated from element $(i-1, j-1), \text{for } i,j \in [1, L-1]$ as
\begin{align}
\mat{\breve{R}}_{i,j}[n] &=  \mat{\breve{R}}_{i-1,j-1}[n]  + \sum_{n'=n-K}^{n+K} (x_{i+N_L-1}^{\Delta}[n']x_{j+N_L-1}^{\Delta}[n']^* \nonumber \\
 &- x_{i-1}^{\Delta}[n']x_{j-1}^{\Delta}[n']^H) \label{eq:sliding}
\end{align}
%The same is true in time where
%\begin{align}
%\mat{\breve{R}}[n] &= \mat{\breve{R}}[n-1] + \sum_{l=0}^{N_L-1} (\vec{x}^{\Delta}_l[n+K]\vec{x}^{\Delta}_l[n+K]^H \nonumber \\
%&- \vec{x}^{\Delta}[n-K-1]\vec{x}^{\Delta}[n-K-1]^H).
%\end{align}
%- \mat{\breve{R}}[n-K-1] + \mat{\breve{R}}[n-K].
Both sliding across subarrays and in time can be realized simultaneously if we hold an $N_K$-average in time of the outer products $\vec{x}^{\Delta}[n]\vec{x}^{\Delta}[n]^H$ in memory and calculate $\mat{\hat{R}}$ averaged over subarrays from this buffer. However, since $N_K$ typically is small, and we have a limited amount of memory per thread on the GPU, we have chosen to focus on sliding across subarrays only (\ref{eq:sliding}). In that case, the complexity of calculating $\mat{\hat{R}}$ is reduced to $O(L^2)$. Important to notice is that sliding will break independence in the sliding dimension. With our way of calculating $\mat{\hat{R}}$, calculations are still independent across time, but not along diagonals in $\mat{\hat{R}}$. We have therefore selected to use one thread per diagonal in the upper triangle of $\mat{\hat{R}}$. And for small subarrays, several covariance matrices are calculated per thread block (Figure \ref{fig:gpulayout}). Maximum supported $L$ is with this approach increased to 512, however for small $L$ high $M$ configurations, memory consumption per thread will be too high to achieve full occupancy of the GPU.

@@

Diagonal loading as a constant value can be added with marginal cost. For the adaptive diagonal loading factor in (\ref{eq:diag}), the trace of $\mat{\hat{R}}$ has to be known before diagonal elements are written back to global memory. Since one row for all covariance matrices is written to global memory per kernel iteration, the diagonal elements need to be held in a local memory until the trace is accumulated. When it is found, scattered writes of the diagonal elements to global memory have to be performed.

@@

\subsection{Compute Beamformer Output}
The formula in (\ref{eq:z_mv}) has been implemented with one thread per element-weight ($L$ threads per weight vector). After the solution vector $\vec{b}$ has been placed in shared memory, one of the $L$ threads finds the sum of $\vec{b}$. Each of the $L$ threads then reads one element of $\vec{b}$ and scales it with the inverse sum just found to get the weight vector $\vec{w}[n]$.

@@

\begin{figure*}[!t]
\centerline{\subfloat[]{\includegraphics[width=3.9in]{gfx/benchmark_us_32_lr.png} \label{fig:benchUS32}}
\hfil
\subfloat[]{\includegraphics[width=3.9in]{gfx/benchmark_us_64_lr.png}%
\label{fig:benchUS64}}}
\caption{Benchmark of GPU Capon beamforming for a cardiac image covering a $70\degree$ sector. Execution time is plotted as a function of subarray length $L \in [1, M/2]$. The plot shows execution times for the three steps listed in Section \ref{sec:meth}: Calculation of covariance matrices in red, solver in green, and the final beamformer sum in blue. For each step, there is one line per choice of temporal averaging, $K \in [0, 1, 2]$. (a) A 32 element array. (b) A 64 element array.}
\label{fig:bench}
\end{figure*}

\section{Benchmarks and Videos}
Figure \ref{fig:bench} shows the results from benchmarking the implementation described in Section \ref{sec:meth} on a Nvidia Quadro 6000 graphics card (14 SMs, 1 TFlops single precision, 6 GB global memory, and 144 GB/s memory bandwidth). The problem sizes are calculated based on 32 and 64 element arrays with center frequency 2.5 MHz. By assuming a system bandwidth of 80\% we need a complex sampling frequency of 2 MHz. If we image down to a depth of 15 cm we get approximately 400 samples per range line. To correctly sample a $70\degree$ sector with 64 and 32 element arrays ($\lambda/2$ pitch) we need at least 80 and 40 beams respectively. The presented execution times do not include transfer of data from CPU-side to the GPU. What is included are all calculations, memory transfers from global to near-core GPU memory, and near core memory access.

In the attached videos one will find a cardiac loop (add ref to video), and a simulated phantom (add ref to videos) processed with different types of settings given to the Capon beamformer. Cardiac data was acquired using a 64 element and 3.5 MHz phased array. Simulations were done using a 64 element 2.5 MHz phased array.

@@

A major part of the total execution time comes from solving systems of linear equations. Further research should therefore focus on making this contribution smaller. This can be done either by implementing faster solvers, or reducing the sample covariance matrix size in a novel way (ref?). When doing such a reduction, including the pre-beamforming described above, one should investigate the loss of contrast and resolution compared with full Capon beamforming.

As for now, we have shown that the Capon beamformer can be implemented in a GPU framework. And if we as in \cite{Chen2011} define real-time as 10 fps, all execution times  in Figure (\ref{fig:bench}) are real time. If we on the other hand define real-time as the time it takes to acquire a given image. The image formed using a 32 and 64 element array has a real time requirement of 125 fps and 62.5 fps respectively. There is some distance to cover before that goal is reached for all values of $L \le M/2$. However we see, by the last definition, that real time is possible for $M=64, L < 16$, and for $M = 32, L < 13$.

(Should comment on videos)

>>>>

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{
Jon Petter \AA{}sen\IEEEauthorrefmark{1},
Jo Inge Buskenes\IEEEauthorrefmark{2},
Carl-Inge C. Nilsen\IEEEauthorrefmark{2},
Andreas Austeng\IEEEauthorrefmark{2},
Sverre Holm\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}
}
\IEEEauthorblockA{
\IEEEauthorrefmark{1} MI-Lab, Norwegian University of Science and Technology, Trondheim, Norway,\\
\IEEEauthorrefmark{2} Department of Informatics, University of Oslo, Oslo, Norway
}
}

@@

For a typical cardiac ultrasound image of $80\times400$ pixels ($70\degree$ sector, 15 cm range) acquired using a 2.5 MHz, M=64 element phased array, we obtain 10 fps (subarray length L=M/2, temporal smoothing over 3 samples). If we do a 2-element pre-beamforming to get the channel count down to 32, the frame rate increases to 44 fps. For a 32 element phased array we need less beams to cover the sector ($40\times400$ pixels), hence with the same parameters the frame rate increases to 87 fps. After optimization the bottleneck of the Capon algorithm is to find the inverse covariance matrix for all pixels. This is an $O(L^3)$ operation, and when L is larger than 16 (M = 64), real time performance is not possible for cardiac imaging with current high-end hardware. The target GPU has been Nvidia’s Quadro 6000, capable of 1 Tflops.

@@

\begin{figure}
\centerline{\includegraphics[width=3in]{gfx/beamforming_mv.pdf}}
\caption{Adaptive beamforming. Impinging signal is demodulated, and aligned in phase by applying steering and focusing delays $\Delta_m[n]$. Adaptive weights, $\vec{w}[n]$, are then calculated based on in-phase data, and finally the output is formed by a weighted coherent sum.}
\label{fig:mvbf}
\end{figure}

@@

Second, we calculate the Capon weight vector as
\begin{align}\label{eq:w}
\vec{w}[n] = \frac{\mat{\hat{R}}[n]^{-1}\vec{a}}{\vec{a}^H\mat{\hat{R}}[n]^{-1}\vec{a}} = \frac{\vec{b}}{\vec{a}^H\vec{b}},
\end{align}
where $\vec{a} = [1_0, 1_1, ..., 1_{L-1}]^T = \vec{1}_L$ when $\vec{x}$ is pre-delayed. 
And last, the beamformer output is calculated by combining all $N_L$ subarrays weighted with the same set of adaptive weights,
\begin{align}
z[n] &= \frac{1}{N_L}\vec{w}[n]^H \sum_{l=0}^{N_L-1} \vec{x}_l^{\Delta}[n]. \label{eq:z_mv}
\end{align}

@@

\subsection{GPU Compute Model}
CUDA by Nvidia has been our choice of GPU developer environment \cite{Nvidia2011}. The CUDA compute model is based on execution of a kernel function across a grid of compute threads, where the kernel function describes the work to be done by each thread. As depicted in Figure \ref{fig:gpulayout}, each position in the grid holds a block of threads, with maximal size of 1024 threads, that supports low-cost synchronization if needed. 
%Each block is further divided into groups of 32 threads, known as a warp. The maximum number of threads inside a block is restricted to 512, hence 16 warps. 
From a CUDA perspective the GPU consists of one or more stream multi processors (SMs), where each SM performs single-instruction-multiple-data (SIMD) operations. A group of blocks is scheduled to each SM until all blocks in the grid have been processed.

An SM has available a limited amount of near-core memory and registers. Each thread therefore has to be as light on memory as possible and has to perform enough instructions per used byte in order to hide memory latency. Global reads and writes on the GPU, in addition to CPU-GPU transfers, have high penalty, and should be minimized as much as possible. All of these constraints combined pose a real challenge as the target problem has to map on to the parallel framework both on a grid, block and memory consumption level in order to benefits from parallel acceleration. In the next section, we will describe how this mapping can be done for the Capon beamformer.

@@

\section{Parallel Capon}\label{sec:meth}
If we analyze (\ref{eq:z_mv}) in combination with (\ref{eq:w}) and (\ref{eq:R}), we see that each calculation of a set of Capon weights is independent across the output $z[n]$. As shown in Figure \ref{fig:gpulayout}, a first level of granularity is therefore to divide the image into a grid, where the output in each cell can be calculated independently of all others. Further, based on memory dependencies and to hold memory consumption per thread at a minimum, one should recognise that a group of threads should process one output in the final image. Continuing with the calculation of one weight vector, the algorithm can be broken down into three main steps that, to some extent, can be further divided into a small set of parallel processes. That is, for each data vector $\vec{x}^\Delta$ in the image:
\begin{align*}
%\begin{array}{l}
\text{1) } &\text{Calculate the sample covariance matrix } \mat{\hat{R}} \text{ as in } \text{(\ref{eq:diag})}\\
\text{2) } &\text{Solve the linear system of equations } \vec{b} =\mat{\hat{R}}^{-1}\vec{1}_L\\
\text{3) } &\text{Calculate } \vec{w} = \vec{b}/\vec{1}_{L}^H\vec{b} \text{ and the amplitude as in (\ref{eq:z_mv})}
%\end{array}
\end{align*}

@@

\subsection{Calculating Multiple Sample Covariance Matrices}
The straightforward sum in (\ref{eq:R}) has complexity $O(L^2N_LN_K)$. Note that for a typical array configuration ($M \in [32, 128]$) we do more spatial than temporal smoothing. Hence, $N_K$ is usually small compared to the number of subarrays ($N_L$). If we let $L = M/\alpha$, $L^2N_L$ becomes $(\alpha - 1)L^3 + L^2$, $\alpha \in [1, M]$. A common choice for $\alpha$ is between two and four \cite{Synnevag2007, Mehdizadeh2012, Synnevag2007a, Austeng2008}. In this interval, calculating the sample covariance matrix is cubic with $L$. This is actually the same as for inverting an $L \times L$ matrix. However little attention has been devoted to this step in the literature.

From (\ref{eq:R}) it is obvious that calculation of one element in $\mat{\hat{R}}$ is independent of all the other elements. A natural level of granularity would then be to assign one thread to each element of $\mat{\hat{R}}$ as in \cite{Chen2011}. This will minimize global reads and writes per thread, but unfortunately the maximum number of threads per block puts an upper limit on the subarray length we then can process, $L_{max}=32$, in order to avoid block-to-block synchronization.

The cubic complexity with $L$ can be reduced by realizing that the calculation in (\ref{eq:R}) overlaps across subarrays. Element $(i,j)$ of $\mat{\hat{R}}[n]$ can be calculated from element $(i-1, j-1), \text{for } i,j \in [1, L-1]$ as
\begin{align}
\mat{\breve{R}}_{i,j}[n] &=  \mat{\breve{R}}_{i-1,j-1}[n]  + \sum_{n'=n-K}^{n+K} (x_{i+N_L-1}^{\Delta}[n']x_{j+N_L-1}^{\Delta}[n']^* \nonumber \\
 &- x_{i-1}^{\Delta}[n']x_{j-1}^{\Delta}[n']^H) \label{eq:sliding}
\end{align}
%The same is true in time where
%\begin{align}
%\mat{\breve{R}}[n] &= \mat{\breve{R}}[n-1] + \sum_{l=0}^{N_L-1} (\vec{x}^{\Delta}_l[n+K]\vec{x}^{\Delta}_l[n+K]^H \nonumber \\
%&- \vec{x}^{\Delta}[n-K-1]\vec{x}^{\Delta}[n-K-1]^H).
%\end{align}
%- \mat{\breve{R}}[n-K-1] + \mat{\breve{R}}[n-K].
%Both sliding across subarrays and in time can be realized simultaneously if we hold an $N_K$-average in time of the outer products $\vec{x}^{\Delta}[n]\vec{x}^{\Delta}[n]^H$ in memory and calculate $\mat{\hat{R}}$ averaged over subarrays from this buffer. However, since $N_K$ typically is small, and we have a limited amount of memory per thread on the GPU, we have chosen to focus on sliding across subarrays only. 
The complexity of calculating $\mat{\hat{R}}$ is now reduced to $O(L^2)$. Important to notice is that sliding will break independence in the sliding dimension. The calculation in (\ref{eq:sliding}) is still independent across time, but not along diagonals in $\mat{\hat{R}}$. We have therefore selected to use one thread per diagonal in the upper triangle of $\mat{\hat{R}}$. And for small subarrays, several covariance matrices are calculated per thread block (Figure \ref{fig:gpulayout}). Maximum supported $L$ is with this approach increased to 512, however for small $L$ high $M$ configurations, memory consumption per thread will be too high to achieve full occupancy of the GPU.

@@

Diagonal loading as a constant value can be added with marginal cost. For the adaptive diagonal loading factor in (\ref{eq:diag}), the trace of $\mat{\hat{R}}$ has to be known before diagonal elements are written back to global memory. Since one row for all covariance matrices is written to global memory per kernel iteration, the diagonal elements need to be held in near-core memory until the trace is accumulated. When it is found, scattered writes of the diagonal elements to global memory have to be performed.

@@

\subsection{Compute Beamformer Output}
The formula in (\ref{eq:z_mv}) has been implemented with one thread per element-weight ($L$ threads per weight vector). After the solution vector $\vec{b}$ has been placed in near-core memory, one of the $L$ threads finds the sum of $\vec{b}$. Each of the $L$ threads then reads one element of $\vec{b}$ and scales it with the inverse sum just found to get the weight vector $\vec{w}[n]$.

@@

\begin{figure*}[!t]
\centerline{\subfloat[]{\includegraphics[width=3.49in]{gfx/benchmark_bar_32_40_400.png} \label{fig:benchUS32}}
\hfil
\subfloat[]{\includegraphics[width=3.49in]{gfx/benchmark_bar_64_80_400.png}%
\label{fig:benchUS64}}}
\caption{Benchmark of GPU Capon beamforming for a cardiac image covering a $70\degree$ sector. Execution time is plotted as a function of subarray length $L \in [1, M/2]$. The figures show execution times for the three steps listed in Section \ref{sec:meth}: Calculation of covariance matrices in light gray, solver in gray, and the final beamformer sum in dark gray. For the calculation of covariance matrices, there is one bar per choice of temporal averaging, $K \in [0, 1, 2]$. (a) A 32 element array. (b) A 64 element array.}
\label{fig:bench}
\end{figure*}

\section{Benchmarks and Videos}
Figure \ref{fig:bench} shows the results from benchmarking the implementation described in Section \ref{sec:meth} on an Nvidia Quadro 6000 graphics card (14 SMs, 1 TFlops single precision, 6 GB global memory, and 144 GB/s memory bandwidth). The problem sizes are calculated based on 32 and 64 element arrays with center frequency 2.5 MHz. By assuming a system bandwidth of 80\% we need a complex sampling frequency of 2 MHz. If we image down to a depth of 15 cm we get approximately 400 samples per range line. To correctly sample a $70\degree$ sector with 64 and 32 element arrays ($\lambda/2$ pitch) we need at least 80 and 40 beams respectively. The presented execution times do not include transfer of data from CPU-side to the GPU. What is included are all calculations, memory transfers from global to near-core GPU memory, and near core memory access.

At (cite) one will find videos of a cardiac loop, 
%(add ref to video)
and a simulated phantom 
%(add ref to videos) 
processed with different types of settings given to the Capon beamformer. Cardiac data was acquired using a 64 element and 3.5 MHz phased array. Simulations were done using a 64 element 2.5 MHz phased array.

@@

A major part of the total execution time comes from solving systems of linear equations. Further research should therefore focus on making this contribution smaller. This can be done either by implementing faster solvers, or reducing the sample covariance matrix size in a novel way \cite{Nilsen2009}. %When doing such a reduction, including the pre-beamforming described above, one should investigate the loss of contrast and resolution compared with full Capon beamforming.
As for now, we have shown that the Capon beamformer can be implemented in a GPU framework. And if we as in \cite{Chen2011} define real-time as 10 fps, all execution times  in Figure (\ref{fig:bench}) are real time. If we on the other hand define real-time as the time it takes to acquire a given image, the image formed using a 32 and 64 element array has a real time requirement of 125 fps and 62.5 fps respectively. There is some distance to cover before that goal is reached for all values of $L \le M/2$. However we see that real time, by the last definition, is possible for $M=64, L < 16$, and for $M = 32, L < 13$. 
(Should comment on videos? Ran out of space)
